The {TestIdeals} package for \emph{Macaulay2}

### 2019-04-29 Fsing/TestIdealsPaper/TestIdeals.tex
Minor changes

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y,z];

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : S = ZZ/5[x,y,z];

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : p = 3;

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i8 : S = ZZ/3[a..f,x,y];

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : R = ZZ/2[a,b,c,d];

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i10 : B = ascendModule(1, A, U)

@@

\begin{remark}[Optimizations in \texttt{ascendIdeal} and other \texttt{testModule} computations]
Throughout the computations described above, we very frequently use the following fact:
\[
(f^p \cdot J)^{[1/p]} = f \cdot (J^{[1/p]}).
\]
To access this enhancement, one should try to pass functions like \texttt{ascendIdeal} and \texttt{frobeniusRoot} the elements and their exponents (see the documentation).
In particular, when computing the $p^e$-th Frobenius root of an ideal of the form $f^n \cdot J$, one should not multiply out $f^n$ and $J$, nor even raise $f$ to the $n^\mathrm{th}$ power, but rather call \texttt{frobeniusRoot(e,n,f,J)}.
\end{remark}

@@

If such a $\Phi_e$ exists, it can be identified to a generator of the module $(I^{[p^e]} : I) / I^{[p^e]}$ by Fedder's Lemma \cite{FedderFPureRat}.  Hence we can find a corresponding\footnote{This is done via the function \texttt{QGorensteinGenerator}. } $u \in I^{[p^e]} : I$.  In this case, if $c \in S$ is the pre-image of a test element of $R$, then setting $I_0 = cR$, it follows that
\[
\tau(R) = {\tt ascendIdeal(e, u, I_0)},
\]
where \texttt{ascendIdeal} is the method explained above, in \autoref{eq.AscendIdealExplanation}.
The $e$ here means all Frobenius roots are taken as multiples of $e$.  In other words, we first compute ${\tt I_0 +  (u \cdot I_0)^{1/p^{[e]}}}$.  Then we compute
\[
{\tt I_0 +  (u\cdot I_0)^{1/p^{[e]}} + (u^{p^e + 1} \cdot I_0)^{1/p^{[2e]}} }
\]
etc.
%
%\pedro{I am not sure what this last sentence is saying}
%

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : T = ZZ/3[x,y];

@@

@@

@@

Again, consider the maps $T^e : \omega_{R^{1/p^e}} \to \omega_R$ we have considered throughout this section.  It is a theorem of Hartshorne-Speiser, Lyubeznik, and Gabber \cite{HartshorneSpeiserLocalCohomologyInCharacteristicP,LyubeznikFModulesApplicationsToLocalCohomology,Gabber.tStruc} that the descending images
\begin{equation}\label{eq: descending images}
\omega_R \supseteq T(\omega_{R^{1/p}}) \supseteq \dots \supseteq T^e(\omega_{R^{1/p^e}}) \supseteq T^{e+1}(\omega_{R^{1/p^{e+1}}}) \supseteq \cdots
\end{equation}
stabilize for $e \gg 0$.
The function \texttt{FPureModule} computes the stable submodule in this chain, called the \emph{HSLG module}, and returns a sequence containing the following items:
\begin{enumerate}[(1)]
   \item The HSLGmodule;
   \item The canonical module, embedded (non-uniquely) as an ideal;
   \item The element $u$ representing the map on the canonical module (see
   \autoref{subsec.DualToFrobeniusOnQuotientRings});
   \item The value of $e > 0$ at which the descending images in \eqref{eq: descending images} stabilize.  This is sometimes also called the HSLG number of the canonical module as a Cartier module.
\end{enumerate}

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : R = ZZ/3[x,y,z]/(x^3 + y^4 + z^5);

@@

                          2        2   3
o3 = ideal (0, y*z, x*z, y , x*y, x , z )

@@

o5 = 1
\end{MyVerbatim}
}\medskip

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : R = ZZ/2[x,y,z];

@@

o3 = ideal (x, y, z)

@@

i4 : u=f^3;

@@

o6 = ideal (x, y, z)

i7 : ((FPureModule(1, f))#3) + 1

o7 : 2
\end{MyVerbatim}
}\medskip

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y,z];

@@

             2        3
o6 = ideal (z , x*z, x )

@@

i5 : frobeniusPower(1/121, ideal f^120)

             2        3
o6 = ideal (z , x*z, x )

i7 : ((FPureModule(1, f))#3) + 1

o7 : 2
\end{MyVerbatim}
}\medskip

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : R = ZZ/3[u,v];

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i1 : R = ZZ/2[a..f];

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i6 : omega = canonicalIdeal S

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i11 : L = compatibleIdeals u;

@@

        3       3       2                         2     2 2       2 2    3 2
ideal (a b*c + a b*d + a c*d*e + a*b*c*d*e + a*b*d e + b d e + c*d e  + d e )

@@

ideal (d, a, e)

@@

@@

ideal (d, b, a)

ideal (c + d, b, a, e)

ideal (b, a, e)
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small\setstretch{.7}
\begin{MyVerbatim}
i13 : print \ unique apply(L, J -> J : omega);

        3       3       2                         2     2 2       2 2    3 2
ideal (a b*c + a b*d + a c*d*e + a*b*c*d*e + a*b*d e + b d e + c*d e  + d e )

@@

@@

ideal (d, b, a)

ideal (e, b, a)
\end{MyVerbatim}
}
\medskip

>>>>

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y,z];

@@

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : S = ZZ/5[x,y,z];

@@

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : p = 3;

@@

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i8 : S = ZZ/3[a..f,x,y];

@@

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/2[a,b,c,d];

@@

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i10 : B = ascendModule(1, A, U)

@@

\begin{remark}[Optimizations in \texttt{ascendIdeal} and other \texttt{testModule} computations]
Throughout the computations described above, we very frequently use the following fact:
\[
(f^p \cdot J)^{[1/p]} = f \cdot (J^{[1/p]}).
\]
To access this enhancement, one should try to pass functions like \texttt{ascendIdeal} and \texttt{frobeniusRoot} the elements and their exponents (see the documentation).
In particular, when computing the $p^e$-th Frobenius root of an ideal of the form $f^n \cdot J$, one should not multiply out $f^n$ and $J$, nor even raise $f$ to the $n^\mathrm{th}$ power, but rather call \texttt{frobeniusRoot(e, n, f, J)}.
\end{remark}

@@

If such a $\Phi_e$ exists, it can be identified to a generator of the module $(I^{[p^e]} : I) / I^{[p^e]}$ by Fedder's Lemma \cite{FedderFPureRat}.  Hence we can find a corresponding\footnote{This is done via the function \texttt{QGorensteinGenerator}. } $u \in I^{[p^e]} : I$.  In this case, if $c \in S$ is the pre-image of a test element of $R$, then setting $I_0 = cR$, it follows that
\[
\tau(R) = {\tt ascendIdeal(e, u, I_0)},
\]
where \texttt{ascendIdeal} is the method explained above, in \autoref{eq.AscendIdealExplanation}.
The $e$ here means all Frobenius roots are taken as multiples of $e$.  In other words, we first compute ${\tt I_0 +  (u \cdot I_0)^{1/p^{[e]}}}$.  Then we compute
\[
{\tt I_0 +  (u\cdot I_0)^{1/p^{[e]}} + (u^{p^e + 1} \cdot I_0)^{1/p^{[2e]}} }
\]
etc.

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : T = ZZ/3[x,y];

@@

o3 : RingMap T <--- S

@@

o7 = ideal 1

o7 : Ideal of R
\end{MyVerbatim}
}\medskip

@@

Again, consider the maps $T^e : \omega_{R^{1/p^e}} \to \omega_R$ we have considered throughout this section.  It is a theorem of Hartshorne-Speiser, Lyubeznik, and Gabber \cite{HartshorneSpeiserLocalCohomologyInCharacteristicP,LyubeznikFModulesApplicationsToLocalCohomology,Gabber.tStruc} that the descending images
\begin{equation}\label{eq: descending images}
\omega_R \supseteq T(\omega_{R^{1/p}}) \supseteq \dots \supseteq T^e(\omega_{R^{1/p^e}}) \supseteq T^{e+1}(\omega_{R^{1/p^{e+1}}}) \supseteq \cdots
\end{equation}
stabilize for $e \gg 0$.
The function \texttt{FPureModule} computes the stable submodule in this chain, called the \emph{HSLG module}, and returns a sequence containing the following items:
\begin{enumerate}[(1)]
   \item The HSLG module;
   \item The canonical module, embedded (non-uniquely) as an ideal;
   \item The element $u$ representing the map on the canonical module (see
   \autoref{subsec.DualToFrobeniusOnQuotientRings});
   \item The value of $e > 0$ at which the descending images in \eqref{eq: descending images} stabilize.  This is sometimes also called the HSLG number of the canonical module as a Cartier module.
\end{enumerate}

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/3[x,y,z]/(x^3 + y^4 + z^5);

@@

                       2        2   3
o3 = ideal (y*z, x*z, y , x*y, x , z )

@@

o5 = 1

i6 : R = ZZ/3[x,y,z]/(x^3 + y^4 + z^5);

i7 : L = FPureModule R;

i8 : L#0

                       2        2   3
o8 = ideal (y*z, x*z, y , x*y, x , z )

i9 : L#1

o9 = ideal 1

o9 : Ideal of R

i10 : L#3

o10 = 1
\end{MyVerbatim}
}\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/2[x,y,z];

@@

o3 = ideal (z, y, x)

@@

i4 : u = f^3;

@@

o5 = ideal (z, y, x)

o5 : Ideal of R

i6 : ((FPureModule(1, f))#3) + 1

o6 = 2
\end{MyVerbatim}
}\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y,z];

@@

             2        3
o3 = ideal (z , x*z, x )

@@

i4 : frobeniusPower(1/121, ideal f^120)

             2        3
o4 = ideal (z , x*z, x )

i5 : ((FPureModule(1, f))#3) + 1

o5 = 2
\end{MyVerbatim}
}\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/3[u,v];

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/2[a..f];

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i6 : omega = canonicalIdeal S

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i11 : L = compatibleIdeals u;

@@

       3       3       2                         2     2 2       2 2    3 2
ideal(a b*c + a b*d + a c*d*e + a*b*c*d*e + a*b*d e + b d e + c*d e  + d e )

@@

ideal (e, a, d)

@@

ideal (e, b, a)

ideal (e, b, a, c + d)

@@

ideal (d, b, a)
\end{MyVerbatim}
}
\medskip

@@

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i13 : print \ unique apply(L, J -> J : omega);

       3       3       2                         2     2 2       2 2    3 2
ideal(a b*c + a b*d + a c*d*e + a*b*c*d*e + a*b*d e + b d e + c*d e  + d e )

@@

ideal (e, b, a)

@@

ideal (d, b, a)
\end{MyVerbatim}
}
\medskip
